---
title: "GRTS GRID comparison"
author: "Shive Lab (B. Baker)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sf)
library(here)
library(janitor)

```


```{r data}

grts <- read_csv(here("outputs/grts_points_easier_groves_table_summary.csv")) %>% 
  clean_names()
grid_summary <- read_csv(here("outputs/grid_points_easier_groves_table_summary.csv")) %>% 
  clean_names()

grove_poly <- read_sf(here("data/spatial_data/groves/all_groves_REexported.shp")) %>% 
  clean_names() %>% 
  filter(str_detect(land_ste_1, "SEKI"))

use_crs <- st_crs(grove_poly) #set global crs

buff_grove_poly <- read_sf(here("data/spatial_data/groves/FINAL_SEGI_AnalysisArea_wOwnerships_GroveNames.shp")) %>% 
  clean_names() %>% 
  st_transform(use_crs)

grts_sf <- read_sf(here("data/spatial_data/grts_grid_strata_spatial_data_9May2025/grts_points_easier_groves_4May2025.shp")) %>% 
  clean_names() %>% 
  st_transform(use_crs)#consistent crs

grid_sf <- read_sf(here("data/spatial_data/grts_grid_strata_spatial_data_9May2025/grid200m_in_needed_strata_easier.shp")) %>% 
  clean_names()%>% 
  st_transform(use_crs)#consisten crs

grove_ranks <- read_csv(here("outputs/groves_summary_rank.csv")) %>% 
  clean_names()

```

```{r}
#remake summary for grts plots
grts_summary <-  grts %>%
  st_drop_geometry() %>%
  select(-c(replsite,stratum,wgt,ip,caty,grove,unit_name)) %>% 
  group_by(strata_nm,grove_name,siteuse) %>%
  summarise(num.grts.plots = length(strata_nm)) %>%
  st_drop_geometry() %>%
  mutate(grov_nm = grove_name, study_design = "grts") %>%
  select(-grov_nm)

```

## to do

1) query the number of BASE plots for grts that are within a buffer of the edge (because most of them seem to be not just clustered, but pushed out to the edge which i think is not representative)
2) decide which groves have a high rank for being visited - ie they have few total plots as well as low plot density (see the rank .csv, but remember you can change how I did the ranking, it was just a first stab)
2) for either approach, i think we assume that we will visit anything that is frontcountry, so make a list excluding those, start looking at where the high priority groves are and how they might be visited with others if they are close in proximity (and then what they offer for strata needs)

##GRTS Base plots and distance from poly edge
```{r}
#set up df
grts_base_sf <- grts_sf %>% 
  filter(siteuse == "Base") %>% #only base plots
  rowid_to_column()

#determine in/out of grove boundary
dist_from_poly <- st_distance(grts_base_sf, grove_poly) #distance from poly

in_grove <- grts_base_sf %>% 
  bind_cols(dist_from_poly) %>% # take distances
  select(!site_id:es_f_ac) %>% # and only distances
  st_drop_geometry() %>% # ignore sf info
  pivot_longer(`...24`:`...62`) %>% # make longer df
  group_by(rowid) %>% # group info for each point together
  slice_min(value) %>% # select closest distance
  mutate(distance_m = round(value), # round to nearest meter
         in_grove = case_when(distance_m == 0 ~ "grove", # if ==0 is within grove
                              T ~ "buffer")) %>% 
  select(rowid, in_grove) # select only needed cols

#Distance from true grove edge
##make polyline
grove_lines <- st_cast(grove_poly, to = "MULTILINESTRING")

grts_base_distances <- grts_base_sf %>% 
  st_distance(grove_lines) #gives distance to each polygon in m

##find closest grove edge to each point
grove_dist <- grts_base_sf %>% 
  bind_cols(grts_base_distances) %>% # take distances
  select(!site_id:es_f_ac) %>% # and only distances
  st_drop_geometry() %>% # ignore sf info
  pivot_longer(`...24`:`...62`) %>% # make longer df
  group_by(rowid) %>% # group info for each point together
  slice_min(value) %>% # select closest distance
  mutate(dist_grv_m = round(value)) %>% # round to nearest meter
  select(rowid, dist_grv_m) # select only needed cols

# #Distance from buffered grove edge
# # some kind of issue here
# ##make polyline
# buff_lines <- st_cast(buff_grove_poly, to = "MULTILINESTRING")
# 
# st_write(buff_lines, here("data/spatial_data/buff_lines_debug.shp"))
# 
# 
# grts_base_distances <- grts_base_sf %>% 
#   st_distance(buff_lines) #gives distance to each polygon in m
# 
# ##find closest grove edge to each point
# buff_dist <- grts_base_sf %>% 
#   bind_cols(grts_base_distances) %>% # take distances
#   select(!site_id:es_f_ac) %>% # and only distances
#   st_drop_geometry() %>% # ignore sf info
#   pivot_longer(`...24`:`...62`) %>% # make longer df
#   group_by(rowid) %>% # group info for each point together
#   slice_min(value) %>% # select closest distance
#   mutate(dis_buff_m = round(value)) %>% # round to nearest meter
#   select(rowid, dis_buff_m) # select only needed cols

grts_base_wDist_sf <- grts_base_sf %>% 
  #left_join(buff_dist) %>% 
  left_join(in_grove) %>% 
  left_join(grove_dist) %>% #join back to rest of info about plots
  mutate(
    dist_cat = case_when(dist_grv_m<10 ~ "within 10m",
                         dist_grv_m<30 ~ "within 30m",
                         dist_grv_m<50 ~ "within 50m",
                         T ~ "50m or further")
  )

#st_write(grts_base_wDist_sf, here("data/spatial_data/grts_base_wDist.shp"))


```

```{r}
plots_total <- grts_base_wDist_sf %>% 
  st_drop_geometry() %>% 
  summarise(.by = grove, total = n())

#summarize distances
base_overall_inGrove <- grts_base_wDist_sf %>% 
  st_drop_geometry() %>% 
  summarise(.by = in_grove, count = n()) %>% 
  mutate(prop = count/257)

base_byGrove_inGrove <- grts_base_wDist_sf %>% 
  st_drop_geometry() %>% 
  summarise(.by = c(grove, in_grove), count = n()) %>% 
  left_join(plots_total) %>% 
  mutate(prop = count/total) %>% 
  select(-total)

base_byGrove_dstBoundary <- grts_base_wDist_sf %>% 
  st_drop_geometry() %>% 
  summarise(.by = c(grove, in_grove, dist_cat), count = n()) %>% 
  left_join(plots_total) %>% 
  mutate(prop = count/total) %>% 
  select(-total) %>% 
  arrange(grove, in_grove)

base_overall_dstBoundary <- grts_base_wDist_sf %>% 
  st_drop_geometry() %>% 
  summarise(.by = c(in_grove, dist_cat), count = n()) %>% 
  mutate(prop = count/257)

```

